State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-06-16 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
edited the spatial frequency results, and spatial freq method to further justify the basis function width

The width parameter of the Gaussian basis functions can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. To demonstrate this, we begin with the two-dimensional Gaussian basis function defined as
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
with the corresponding Fourier transform
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{cy}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent the field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

@@

Given the basis function separation, the basis function width was set to $\sigma_{\phi}^2=2.5$. This facilitated a smooth representation of the field by the basis function decomposition given the basis function separation and provided further low-pass spatial filtering of the disturbance. From equation~\ref{eq:WidthCutOffRelationship}, this allowed the basis function decomposition to represent spatial frequencies up to approximately $0.12$~Hz. The spatial frequency response of the basis functions can be seen in Figure~\ref{fig:BasisFunctionFreqResponse}.
\begin{figure} 
	\begin{center}
		\subfigure[][]{\label{fig:ObservationFreqAnal}\includegraphics{./Graph/ObservationFrequencyResponse.pdf}}\\	
		\subfigure[][]{\label{fig:SensorFunctionFreqResponse}\includegraphics{./Graph/SensorChar.pdf}}
		\subfigure[][]{\label{fig:BasisFunctionFreqResponse}\includegraphics{./Graph/BasisChar.pdf}}
	\end{center}
	\caption{Spatial frequency analysis. \textbf{a.} Two 2-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field. The symmetry of the sensors and field basis functions allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b.} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} The spatial frequency response of a 1-dimensional neural field basis function.} 
\end{figure}

>>>>

The width parameter of the Gaussian basis functions can also be inferred using spectral considerations~\cite{Sanner1992,Scerri2009}. Alternatively, the upper spatial frequency content of the neural field modelled by the basis function decomposition can be determined, given the basis function width. Since Shannon's sampling theorem is derived for perfect point sensors, this is particularly useful when considering spatially distributed basis functions. To demonstrate this, we begin with the two-dimensional Gaussian basis function defined as
\begin{equation}
 \phi(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_{\phi}^2} \mathbf r^\top\mathbf r}\right)
\end{equation}
with the corresponding Fourier transform
\begin{equation}\label{eq:GaussianFT}
\boldsymbol\Phi(\boldsymbol \nu)=\left(\frac{1}{\pi\sigma_{\nu}^2}\right)\mathrm{exp}\left(-\frac{1}{\sigma_{\nu}^2}\boldsymbol\nu^\top \boldsymbol\nu\right),
\end{equation}
where 
\begin{equation}\label{eq:GaussianFTWidth}
	\sigma^2_{\nu} = \frac{1}{\pi^2\sigma_{\phi}^2}. 
\end{equation}
The basis function width, $\sigma^2_{\phi}$, should be chosen to give at least 3~dB attenuation at $\boldsymbol\nu_{cy}$ giving
\begin{equation}\label{eq:WidthCutOffRelationship}
 \sigma^2_{\phi}= \frac{1}{\pi^2\sigma_{\nu_{cy}}^2},
\end{equation}
where
\begin{equation}\label{eq:WidthFrequencyRelationship}
 \sigma^2_{\nu_{cy}}= \frac{2\boldsymbol\nu_{cy}^\top \boldsymbol\nu_{cy}}{\ln 2}.
\end{equation}
This ensures that the basis functions can represent the field with frequency content up to $\boldsymbol\nu_{cy}$. Derivations for equations~\ref{eq:GaussianFT} and \ref{eq:WidthFrequencyRelationship} are given in~\ref{ap:FrequencyAnalysis}.

@@

Given the basis function separation, the basis function width was set to $\sigma_{\phi}^2=2.5$~mm. This facilitated a smooth representation of the field by the basis function decomposition given the basis function separation and provided low-pass spatial filtering of the field. From equation~\ref{eq:WidthCutOffRelationship}, this allowed the basis function decomposition to represent spatial frequencies up to approximately $0.12$~Hz. The spatial frequency response of the basis functions can be seen in Figure~\ref{fig:BasisFunctionFreqResponse}.
\begin{figure} 
	\begin{center}
		\subfigure[][]{\label{fig:ObservationFreqAnal}\includegraphics{./Graph/ObservationFrequencyResponse.pdf}}\\	
		\subfigure[][]{\label{fig:SensorFunctionFreqResponse}\includegraphics{./Graph/SensorChar.pdf}}
		\subfigure[][]{\label{fig:BasisFunctionFreqResponse}\includegraphics{./Graph/BasisChar.pdf}}
	\end{center}
	\caption{Spatial frequency analysis. \textbf{a.} Two 2-dimensional spatial frequency (in power dB, averaged over 400 time steps) of the observed field. The symmetry of the sensors and field basis functions allow for the 1-dimensional representation of the spectral characteristics. The dashed line shows the cut-off frequency (-3 dB point). \textbf{b.} The spatial frequency response of a 1-dimensional model sensor. The dashed line shows the cut-off frequency (-3 dB point), indicating a wider bandwidth than the field basis function. \textbf{c.} The spatial frequency response of a 1-dimensional neural field basis function.} 
\end{figure}
